```{r}
#| include: false
#| echo: false
# load packages
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
library(stringr)
library(tidyr)
library(patchwork)
library(gtsummary)
library(cowplot)
library(gt)
library(readxl)
library(knitr)
library(kableExtra)
library(lme4)
library(glmmTMB)
library(marginaleffects)
library(broom.mixed)
library(officer)
library(flextable)
```

## Cases

```{r, fig.width = 9, fig.height=5}
#| echo: false
#| warning: false

#upload db final database -> db_analysis_20may2026.csv

#outbreak sizes by home and unit
outbreak_size <- db_final %>%
  filter(outbreak_size!=1) %>% #remove those with only one case
  group_by(home, unit, virus, outbreak_num, allocation_f, beds,
           outbreak_start, outbreak_end) %>%
  mutate(timing = as.numeric(difftime(case_symp_onset_f,
                                      outbreak_start,
                                      units = "days")),
         suspect_case = replace_na(suspect_case, 0)) %>%
  dplyr::summarize(outbreak_size=n(),
            outbreak_size_secondary=sum(timing>0), 
            outbreak_size_confirmed=sum(suspect_case==0),
            outbreak_size_secondary_confirmed = sum(suspect_case == 0 & timing > 0),
            outbreak_size_secondary_suspect=sum(suspect_case==1 & timing >0)) %>%
  group_by(home) %>%
  mutate(total_outbreaks=n(),
         home_outbreak_number=row_number()) %>%
  ungroup() 

#outbreak duration in days
outbreak_size_y <- outbreak_size %>%
  filter(outbreak_size!=1) %>% #remove those with only one case
  mutate(outbreak_duration = as.numeric(outbreak_end-outbreak_start))

#find overlapping outbreaks
outbreak_size_y_overlap <- outbreak_size_y %>%
  arrange(home, unit, outbreak_start) %>%
  group_by(home, unit) %>%
  mutate(
    # Get lag and lead start/end
    prev_start = lag(outbreak_start),
    prev_end = lag(outbreak_end),
    next_start = lead(outbreak_start),
    next_end = lead(outbreak_end),

    # Does current outbreak start during previous one?
    overlaps_prev = !is.na(prev_start) & outbreak_start >= prev_start & outbreak_start <= prev_end,

    # Does previous outbreak overlap with current? (so we can flag the previous row too)
    overlaps_next = !is.na(next_start) & next_start >= outbreak_start & next_start <= outbreak_end,

    # Flag if this outbreak overlaps with another (either way)
    outbreak_overlap_flag = overlaps_prev | overlaps_next
  ) %>%
  ungroup()

#make table
# Normalize per 100 beds
outbreak_norm <- outbreak_size_y %>%
  mutate(allocation_beds = case_when(allocation_f=="Intervention" ~ 2113,
                                    allocation_f=="Control" ~ 1850)) %>%
  mutate(
    sec_per_100   = (outbreak_size_secondary / allocation_beds) * 100,
    conf_per_100  = (outbreak_size_secondary_confirmed / allocation_beds) * 100,
    susp_per_100  = (outbreak_size_secondary_suspect / allocation_beds) * 100,
    total_per_100   = (outbreak_size / allocation_beds) * 100,
    total_conf_per_100  = (outbreak_size_confirmed / allocation_beds) * 100,
    total_susp_per_100  = (outbreak_size_secondary_suspect / allocation_beds) * 100,
    
  )

# Summary function with custom labels
summarize_metric <- function(df, metric_col, label) {
  overall <- df %>%
    group_by(allocation_f) %>%
    dplyr::summarize(value = sum(.data[[metric_col]], na.rm = TRUE), .groups = "drop") %>%
    mutate(Cases = "  Overall", Section = label)   

  by_virus <- df %>%
    filter(virus %in% c("COVID", "FLU", "RSV")) %>%
    mutate(
      virus_label = case_when(
        virus == "COVID" ~ "SARS-CoV-2",
        virus == "FLU"   ~ "Influenza",
        virus == "RSV"   ~ "RSV"
      )
    ) %>%
    group_by(virus_label, allocation_f) %>%
    dplyr::summarize(value = sum(.data[[metric_col]], na.rm = TRUE), .groups = "drop") %>%
    rename(Cases = virus_label) %>%                     
    mutate(
      Cases = paste0("  ", Cases),                      
      Section = label
    )

  bind_rows(overall, by_virus)
}

# Create summaries
total_summary     <- summarize_metric(outbreak_norm, "total_per_100", "TOTAL")
confirmed_summary <- summarize_metric(outbreak_norm, "total_conf_per_100", "CONFIRMED")
suspect_summary   <- summarize_metric(outbreak_norm, "total_susp_per_100", "SUSPECT")

# Combine all
full_summary <- bind_rows(total_summary, confirmed_summary, suspect_summary)

# Set desired row order
cases_levels <- c("  Overall", "  SARS-CoV-2", "  Influenza", "  RSV")  # renamed variable

# Final table preparation
summary_clean <- full_summary %>%
  pivot_wider(names_from = allocation_f,
              values_from = value,
              names_glue = "{allocation_f}") %>%
  select(Section, Cases, Control, Intervention) %>%    
  mutate(
    Cases = factor(Cases, levels = cases_levels),      
    Section = factor(Section, levels = c("TOTAL", "CONFIRMED", "SUSPECT"))
  ) %>%
  arrange(Section, Cases) %>%                          
  mutate(across(where(is.numeric), ~ round(.x, 1)))

# Final formatted table
summary_clean %>%
  select(Cases, Control, Intervention) %>%             
  kable("latex", booktabs = TRUE, align = "lcc",
        col.names = c("Cases", "Control", "Intervention"),   
        caption = "Cases per 100 Beds by Allocation Group") %>%
  kable_styling(latex_options = c("hold_position", "striped", "scale_down")) %>%
  pack_rows("TOTAL", 1, 4, bold = TRUE, italic = FALSE) %>%
  pack_rows("CONFIRMED", 5, 8, bold = TRUE, italic = FALSE) %>%
  pack_rows("SUSPECT", 9, 12, bold = TRUE, italic = FALSE)
```

\newpage

## Outbreak Sizes

```{r table_summary, results='asis'}
#| echo: false
#| warning: false

#outbreak duration in days
outbreak_size_y <- outbreak_size %>%
  filter(outbreak_size!=1) %>% #remove those with only one case
  mutate(outbreak_duration = as.numeric(outbreak_end-outbreak_start),
         outbreak_size_initial=as.integer(outbreak_size-outbreak_size_secondary),
         outbreak_size_initial_confirmed=as.integer(outbreak_size_confirmed-outbreak_size_secondary_confirmed))

#table
outbreak_size_y %>% 
  select(virus, allocation_f, outbreak_size, outbreak_size_initial,
         outbreak_size_secondary, outbreak_size_confirmed,
         outbreak_size_initial_confirmed,
         outbreak_size_secondary_confirmed, outbreak_duration,) %>%
  tbl_summary(
    by = allocation_f,
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    type = list(outbreak_size_confirmed ~ "continuous",
                outbreak_size_initial ~ "continuous",
                outbreak_size_initial_confirmed ~ "continuous"),
    digits = all_continuous() ~ 2,
    missing_text = "(Missing)",
    label=c(virus ~ "Virus",
            outbreak_size ~ "Total Outbreak Size",
            outbreak_size_confirmed ~ "Total Confirmed Outbreak Size",
            outbreak_size_initial ~ "Initial Cases Per Outbreak",
            outbreak_size_secondary ~ "Secondary Cases Per Outbreak",
            outbreak_size_initial_confirmed ~ "Initial Confirmed Cases Per Outbreak",
            outbreak_size_secondary_confirmed ~ "Secondary Confirmed Cases Per Outbreak",
            outbreak_duration ~ "Outbreak Duration (days)")
    ) %>%
  add_overall() %>%
  as_kable_extra(booktabs = TRUE)
# as_gt() %>% gtsave(filename = "ae_table3.png") # to save output
```

```{r summary-table, results='asis', echo=FALSE}
#| echo: false
#| warning: false

library(flextable)

#add in by virus
x <- outbreak_size %>% 
  filter(outbreak_size!=1) %>% #remove those with only one case
  mutate(outbreak_duration = as.numeric(outbreak_end-outbreak_start),
         outbreak_size_initial=as.integer(outbreak_size-outbreak_size_secondary),
         outbreak_size_initial_confirmed=as.integer(outbreak_size_confirmed-outbreak_size_secondary_confirmed)) %>% 
  select(virus, allocation_f, outbreak_size, outbreak_size_initial,
         outbreak_size_secondary, 
         outbreak_size_confirmed, outbreak_size_initial_confirmed,
         outbreak_size_secondary_confirmed, outbreak_duration) %>%
  tbl_strata(
    strata = virus,
    .tbl_fun =
      ~ .x |>
      tbl_summary(by = allocation_f, 
                  type = c(outbreak_size ~ "continuous",
                           outbreak_duration ~ "continuous",
                           outbreak_size_secondary ~"continuous",
                           outbreak_size_initial ~ "continuous",
                           outbreak_size_secondary_confirmed ~"continuous",
                           outbreak_size_initial_confirmed ~ "continuous",
                           outbreak_size_confirmed ~ "continuous"),
                  statistic = list(
                    all_continuous() ~ "{median} ({p25}, {p75})",
                    all_categorical() ~ "{n} / {N} ({p}%)"
                  ),
                  digits = all_continuous() ~ 0,
                  missing_text = "(Missing)",
                  label=c(outbreak_size ~ "Total Outbreak Size",
                          outbreak_size_initial ~ "Initial Cases Per Outbreak",
                          outbreak_size_secondary ~ "Secondary Cases Outbreak Size",
                          outbreak_size_initial_confirmed ~ "Initial Confirmed Cases Per Outbreak",
            outbreak_size_secondary_confirmed ~ "Secondary Confirmed Cases Per Outbreak",
                          outbreak_duration ~ "Outbreak Duration (days)",
            outbreak_size_confirmed ~ "Total Confirmed Outbreak Size")
      ) 
  ) %>% 
  as_flex_table() %>%
  autofit() %>%
  fontsize(size = 8) %>%
  width(width = 0.9)  # Adjust the width as needed to prevent cutoff 
  #gtsave(filename = "table3.png") # to save output
```

\newpage

## Secondary Outcomes

The secondary outcomes were the outcomes of each resident infected with a virus categorized into recovered at home, transfer to an acute care institution or death. And then separated by virus.

```{r}
#| echo: false
#| warning: false

#case outcomes by allocation arm
db_final %>%
  select(allocation_f, virus, case_outcome, combined_outcome, suspect_case) %>%
  mutate(combined_outcome = case_when(case_outcome==1 ~ "Recovered",
                                      case_outcome==2 ~ "Transfer to hospital",
                                      TRUE~combined_outcome)) %>%
  mutate(case_outcome = case_when(case_outcome==1~"Recovered",
                                  case_outcome==2~"Transfer",
                                  case_outcome==3~"Died",
                                  TRUE~NA)) %>%
  mutate(case_outcome=factor(case_outcome, 
                             levels=c("Recovered",
                                      "Transfer",
                                      "Died"))) %>%
  mutate(combined_outcome_f = case_when(combined_outcome=="Deceased within facility" &
                                          suspect_case==1 ~ "Deceased within facility - Suspect", 
                                        combined_outcome=="Transfer and deceased"&
                                          suspect_case==1 ~ "Transfer and deceased - Suspect",
                                        combined_outcome=="Transfer to hospital" &
                                          suspect_case==1 ~ "Transfer to hospital - Suspect",
                                        TRUE ~ combined_outcome)
  ) %>%
  mutate(combined_outcome_f=factor(combined_outcome_f, 
                             levels=c("Recovered",
                                      "Transfer to hospital",
                                      "Transfer to hospital - Suspect",
                                      "Transfer and deceased",
                                      "Transfer and deceased - Suspect",
                                      "Deceased within facility",
                                      "Deceased within facility - Suspect"))) %>%
  tbl_summary(
    by = allocation_f,
    include=c(virus, combined_outcome_f), #change what to include
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing_text = "(Missing)",
    label=c(combined_outcome_f ~ "Outcomes", #change what to include
            virus~"Virus")
  ) %>%
  add_overall() %>%
   as_kable_extra(booktabs = TRUE)

#outcomes by allocation by virus
db_final %>%
  #group_by(allocation_f, virus, case_outcome) %>%
  #summarise(n=n()) %>%
  #ungroup() %>% 
  select(allocation_f, virus, case_outcome, combined_outcome, suspect_case) %>%
  mutate(combined_outcome = case_when(case_outcome==1 ~ "Recovered",
                                      case_outcome==2 ~ "Transfer to hospital",
                                      TRUE~combined_outcome)) %>%
  mutate(case_outcome = case_when(case_outcome==1~"Recovered",
                                  case_outcome==2~"Transfer",
                                  case_outcome==3~"Died",
                                  TRUE~NA)) %>%
  mutate(case_outcome=factor(case_outcome, 
                             levels=c("Recovered",
                                      "Transfer",
                                      "Died"))) %>%
  mutate(combined_outcome_f = case_when(combined_outcome=="Deceased within facility" &
                                          suspect_case==1 ~ "Deceased within facility - Suspect", 
                                        combined_outcome=="Transfer and deceased"&
                                          suspect_case==1 ~ "Transfer and deceased - Suspect",
                                        combined_outcome=="Transfer to hospital" &
                                          suspect_case==1 ~ "Transfer to hospital - Suspect",
                                        TRUE ~ combined_outcome)
  ) %>%
  mutate(combined_outcome_f=factor(combined_outcome_f, 
                             levels=c("Recovered",
                                      "Transfer to hospital",
                                      "Transfer to hospital - Suspect",
                                      "Transfer and deceased",
                                      "Transfer and deceased - Suspect",
                                      "Deceased within facility",
                                      "Deceased within facility - Suspect"))) %>%
  tbl_strata(
    strata = virus,
    .tbl_fun =
      ~ .x |>
      tbl_summary(by = allocation_f, 
                  include=c(combined_outcome_f), #change what to include
                  statistic = list(
                    all_continuous() ~ "{median} ({p25}, {p75})",
                    all_categorical() ~ "{n} / {N} ({p}%)"
                  ),
                  digits = all_continuous() ~ 2,
                  missing_text = "(Missing)",
                  label=c(combined_outcome_f ~ "Outcomes") #change what to include here
      )
  ) %>%
  as_flex_table() %>%
  autofit() %>%
  fontsize(size = 8) %>%
  width(width = 0.9)  # Adjust the width as needed to prevent cutoff 
  #gtsave(filename = "table3.png") # to save output
```

\newpage

Time from symptom onset to treatment.

```{r}
#| echo: false
#| warning: false

##### time to treatment
t_treatment <- db_final %>%
  filter(is.na(suspect_case)) %>%
  select(study_id, allocation_f, treatment_f, treatment_start_date, case_symp_onset_f, virus) %>%
  mutate(symp_treat_time = as.numeric(
           difftime(as.POSIXct(treatment_start_date), as.POSIXct(case_symp_onset_f), units = "days")
         ))

#summarize treatment in a table
# Create summary table
t_treatment %>%
  tbl_summary(
    by = allocation_f,
    include = c(treatment_f, symp_treat_time),
    missing='no',
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})", 
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 1,
    label = list(
      treatment_f ~ "Received Treatment",
      symp_treat_time ~ "Time to Treatment (days)"
    )
  ) %>%
  add_overall() %>%
  as_kable_extra(booktabs = TRUE)

#### table of overall and separated by virus
# Create summary tables
tbl_overall <- t_treatment %>%
  tbl_summary(
    by = allocation_f,
    include = c(treatment_f, symp_treat_time),
    missing = 'no',
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 1,
    label = list(
      treatment_f ~ "Received Treatment",
      symp_treat_time ~ "Time to Treatment (days)"
    )
  ) %>%
  add_overall()

tbl_covid <- t_treatment %>%
  filter(virus == "COVID") %>%
  tbl_summary(
    by = allocation_f,
    include = c(treatment_f, symp_treat_time),
    missing = "no",
    type = list(
      symp_treat_time = "continuous"
    ),
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 1,
    label = list(
      treatment_f = "Received Treatment",
      symp_treat_time = "Time to Treatment (days)"
    )
  ) %>%
  add_overall()


tbl_flu <- t_treatment %>%
  filter(virus == "FLU") %>%
  tbl_summary(
    by = allocation_f,
    include = c(treatment_f, symp_treat_time),
    missing = 'no',
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 1,
    label = list(
      treatment_f ~ "Received Treatment",
      symp_treat_time ~ "Time to Treatment (days)"
    )
  ) %>%
  add_overall()

# Stack with headers
tbl_stack(
  list(tbl_overall, tbl_covid, tbl_flu),
  group_header = c("Overall", "COVID", "FLU")
) %>%
  as_kable_extra(booktabs = TRUE)

#graph time from symptom onset to treatment
treat_1 <- t_treatment %>%
  mutate(virus = case_when(virus== "COVID" ~ "SARS-CoV-2",
                           virus== "RSV" ~ "RSV",
                           virus== "FLU" ~ "Influenza",
                           TRUE ~ "NACHO"),
         virus = factor(virus, levels = c("SARS-CoV-2", "Influenza", "RSV"))) %>%
  filter(virus!="RSV") %>%
  ggplot(aes(x=allocation_f, y=symp_treat_time)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(width = 0.2) +
  theme_classic() + 
  labs(y="Days from Symptom Onset to Treatment",
       x="Allocation") 
treat_2 <- t_treatment %>%
  mutate(virus = case_when(virus== "COVID" ~ "SARS-CoV-2",
                           virus== "RSV" ~ "RSV",
                           virus== "FLU" ~ "Influenza",
                           TRUE ~ "NACHO"),
         virus = factor(virus, levels = c("SARS-CoV-2", "Influenza", "RSV"))) %>%
  filter(virus!="RSV") %>%
  ggplot(aes(x=allocation_f, y=symp_treat_time)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(width = 0.2) +
  theme_classic() + 
  labs(y="Days from Symptom Onset to Treatment",
       x="Allocation") +
  facet_grid(~virus)

time_treatment <- treat_1 + treat_2 + plot_layout(widths = c(1, 2)) + plot_annotation(tag_levels = 'A')
time_treatment

#save peg
ggsave(time_treatment, filename = "treatment_time.jpeg", width = 8, # width in inches
  height = 6,        # height in inches
  dpi = 300)

#time to oseltamivir treatment start for flu
time_tr_flu <- glm(symp_treat_time ~ allocation_f, 
               family="gaussian",
               data=subset(t_treatment, c(virus=="FLU" & treatment_f=="Yes")))
summary(time_tr_flu)
avg_predictions(time_tr_flu, by = "allocation_f")
avg_comparisons(time_tr_flu, variables = "allocation_f")

#time to oseltamivir treatment start for covid
time_tr_covid <- glm(symp_treat_time ~ allocation_f, 
               family="gaussian",
               data=subset(t_treatment, c(virus=="COVID" & treatment_f=="Yes")))
summary(time_tr_covid)
avg_predictions(time_tr_covid, by = "allocation_f")
avg_comparisons(time_tr_covid, variables = "allocation_f")

#time to antiviral treatment start fo
time_tr_all <- glm(symp_treat_time ~ allocation_f + virus, 
               family="gaussian",
               data=subset(t_treatment, treatment_f=="Yes"))
summary(time_tr_all)
avg_predictions(time_tr_all, by = "allocation_f")
avg_comparisons(time_tr_all, variables = "allocation_f")
```

\newpage

```{r}
#| echo: false
#| warning: false
#| include: false

##number of outbreaks per unit per home by virus
outbreak_num <- outbreak_size %>%
  filter(outbreak_size!=1) %>% #remove those with only one case
  group_by(home, unit, virus, allocation_f) %>%
  count() %>%
  group_by(home, unit, virus, allocation_f) %>%
  dplyr::summarize(n=n()) 

#add in virus for each home so all have flu/covid/rsv
#remove the missing and non-covid/rsv/flu viruses
outbreak_num <- right_join(outbreak_num, 
                           expand.grid(home=toupper(letters[1:20]), 
                                       virus=c("COVID","FLU","RSV"))) %>%
  select(-allocation_f) %>%
  mutate(n=replace_na(n,0))

#load randomization database
library(readxl)
random <- read_excel("prompt_ltc_home_unit_key_scrambled_07april2025.xlsx") %>%
  select(-allocation_f) %>%
  right_join(read_excel("randomization_key_08jan2025.xlsx") %>%
  mutate(allocation_f = case_when(allocation=="Platform" ~ "Intervention",
                                  allocation=="Control" ~ "Control")))

#add in all viruses for all units in each home
out_num <- random %>% 
  cross_join(tibble(virus=c("COVID","FLU","RSV"))) %>%
  left_join(outbreak_num) %>%
  mutate(n=replace_na(n,0),
         home_virus = paste(home, virus))

#outbreak numbers model
library(lme4)
m1 <- glmer(n ~ allocation_f + virus +
                   (1|home), 
                 data=out_num, family = "poisson")
summary(m1)

#outbreak size model
outbreak_size_model <- outbreak_size %>%
  filter(outbreak_size!=1) %>% #remove those with only one case
  mutate(outbreak_size_initial = outbreak_size-outbreak_size_secondary, #case number on day 1
         noncases = beds-outbreak_size_initial, #susceptible cases on each unit
         home_virus = paste(home, virus)) %>% #combine for one variable of home/unit if needed
  ungroup() %>%
  group_by(home, unit, virus) %>%
  mutate(outbreak_virus_num = row_number())

#outbreak size model with number of prior cases being used to create non-cases
#now set up model database
outbreak_size_model_y <- outbreak_size %>%
  filter(outbreak_size != 1) %>%  # Remove single-case outbreaks
  mutate(
    home_virus = paste(home, virus),
    outbreak_size_initial = outbreak_size-outbreak_size_secondary #case number on day 1
  ) %>%
  group_by(home, unit, virus) %>%
  arrange(home, unit, virus, outbreak_start) %>%  # Ensure correct order
  mutate(
    outbreak_virus_num = row_number(),
    prior_outbreak_size_total = lag(cumsum(outbreak_size), default = 0),
    noncases = beds - prior_outbreak_size_total - outbreak_size
  ) %>%
  ungroup()
    
#model outbreak size (secondary cases)  
m2 <- glmer (cbind(outbreak_size_secondary, noncases) ~ 
               allocation_f + virus + (1|home), 
             family="binomial",
             data=outbreak_size_model_y)
summary(m2)

#combine models
library(metafor)
model_combo <- rma(yi=c(fixef(m1)[2], fixef(m2)[2]),
    vi = c(vcov(m1)[2,2], vcov(m2)[2,2]),
    method="FE")

#model proportion of outbreak that are secondary cases
m3 <- glmer(
  cbind(outbreak_size_secondary, outbreak_size - outbreak_size_secondary) ~ allocation_f + 
    virus + (1|home),
  family = "binomial",
  data = outbreak_size_model_y
)
```

```{r abs-risk-death, echo=FALSE, message=FALSE, warning=FALSE}

#deaths per outbreak
deaths_ob <- db_final %>% 
  filter(outbreak_size!=1) %>%
  group_by(home, unit, virus, outbreak_num, allocation_f, beds,
           outbreak_start, outbreak_end) %>%
  mutate(timing = as.numeric(difftime(case_symp_onset_f,
                                      outbreak_start,
                                      units = "days")),
         suspect_case = replace_na(suspect_case, 0)) %>%
  #filter(timing>0) %>% #include or exclude initial cases
  mutate(rowid = row_number(),
         death= case_when(case_outcome==3 ~ 1,
                          TRUE ~0),
         death_14d=as.factor(case_when(as.numeric(difftime(case_death_date,
                                      case_symp_onset_f,
                                      units = "days"))<14 ~ "death14",
                             TRUE ~"no_death")))
#model deaths per outbreak among cases
model_death_logistic_agg <- glmer(
  death ~ allocation_f + virus + (1 | home),
  family = binomial(link = "logit"),
  data = deaths_ob
)
#model deaths within 14d per outbreak among cases
model_death_logistic_agg_14d <- glmmTMB(
  death_14d ~ allocation_f + virus + (1 | home),
  family = binomial(link = "logit"),
  data = deaths_ob
)

#confirmed deaths (ie exclusion of suspect cases)
model_death_logistic_agg_c <- glmer(
  death ~ allocation_f + virus + (1 | home),
  family = binomial(link = "logit"),
  data = subset(deaths_ob, suspect_case==0)
)
```

```{r abs-risk-transfer, echo=FALSE, message=FALSE, warning=FALSE}

#transfers per outbreak
transfers_ob <- db_final %>% 
  filter(outbreak_size!=1) %>%
  group_by(home, unit, virus, outbreak_num, allocation_f, beds,
           outbreak_start, outbreak_end) %>%
  mutate(timing = as.numeric(difftime(case_symp_onset_f,
                                      outbreak_start,
                                      units = "days")),
         suspect_case = replace_na(suspect_case, 0)) %>%
  #filter(timing>0) %>% #include or exclude initial cases
  mutate(rowid = row_number(),
         transfer= case_when(case_outcome==2 ~ 1,
                          TRUE ~0),
         transfer_14d=as.factor(case_when(as.numeric(difftime(case_hosp_transfer_date,
                                      case_symp_onset_f,
                                      units = "days"))<14 ~ "transfer14",
                             TRUE ~"no_transfer")))
#model transfers per outbreak per number of secondary cases
model_transfer_logistic_agg <- glmmTMB(
  transfer ~ allocation_f + virus + (1 | home),
  family = binomial(link = "logit"),
  data = transfers_ob
)
#model transfers per outbreak per number of secondary cases
model_transfer_logistic_agg_14d <- glmmTMB(
  transfer_14d ~ allocation_f + virus + (1 | home),
  family = binomial(link = "logit"),
  data = transfers_ob
)

#model transfers per outbreak per number of secondary cases for confirmed cases only
model_transfer_logistic_agg_c <- glmmTMB(
  transfer ~ allocation_f + virus + (1 | home),
  family = binomial(link = "logit"),
  data = subset(transfers_ob, suspect_case==0)
)
```

\newpage

## Model Outcomes

```{r, results='asis'}
#| echo: false
#| warning: false

# Source the helpers
source("model_summary_helpers.R")

# Define models with their outcome labels
binary_models <- list(
  "Death" = model_death_logistic_agg,
  "Death - Confirmed" = model_death_logistic_agg_c,
  "Transfer" = model_transfer_logistic_agg,
  "Transfer - Confirmed" = model_transfer_logistic_agg_c,
  "Death - 14 Days" = model_death_logistic_agg_14d,
  "Transfer - 14 Days" = model_transfer_logistic_agg_14d,
  "Secondary Transmission" = m2,
  "Proportion of Secondary Cases per Outbreak" = m3,
  "Secondary Transmission - Confirmed" = m2_c,
  "Proportion of Secondary Cases per Outbreak - Confirmed" = m3_c
)

# Define count models with newdata for predictions
count_models <- list(
  list(model = m1, label = "Outbreaks", newdata = datagrid(model = m1, allocation_f = c("Control", "Intervention"), virus = "COVID"))
)

# Create summary for all binary models
binary_summaries <- lapply(names(binary_models), function(label) {
  process_binary_model(binary_models[[label]], label)
})

# Create summary for all count models
count_summaries <- lapply(count_models, function(x) {
  process_count_model(x$model, x$label, newdata = x$newdata)
})

#create summary for outbreak size
outbreak_models <- list(
  list(model = m2, label = "Predicted Outbreak Size", newdata = outbreak_size_model_y),
  list(model = m2_c, label = "Predicted Outbreak Size - Confirmed", newdata = outbreak_size_model_y)
)
#run models
outbreak_summaries <- lapply(outbreak_models, function(x) {
  process_outbreak_size_model(
    model = x$model,
    data = x$newdata,
    outcome_label = x$label
  )
})

# Combine all results
final_summary_table <- bind_rows(count_summaries, binary_summaries) 
final_summary_table <- bind_rows(final_summary_table, outbreak_summaries) 

#order
final_summary_table <- final_summary_table %>%
  mutate(Outcome = factor(Outcome, levels = c("Death", "Death - Confirmed",
                                              "Transfer", "Transfer - Confirmed",
                                              "Outbreaks", 
                                              "Secondary Transmission",
                                              "Secondary Transmission - Confirmed",
                                              "Proportion of Secondary Cases per Outbreak",
                                              "Proportion of Secondary Cases per Outbreak - Confirmed",
                                              "Death - 14 Days",
                                              "Transfer - 14 Days",
                                              "Predicted Outbreak Size",
                                              "Predicted Outbreak Size - Confirmed"))) %>%
  arrange(Outcome)

#make nice with gt
final <- final_summary_table %>%
  select(
    Outcome,
    Control,
    Treatment,
    `Risk Difference`,
    OR_CI,
    p_value
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Outcome Summary by Allocation Group**")
  ) %>%
  cols_label(
    Outcome = "Outcome",
    Control = "Control",
    Treatment = "Intervention",
    `Risk Difference` = "Absolute Difference (Intervention − Control, 95% CI)",
    OR_CI = "Odds Ratio (95% CI)",
    p_value = "p-value"
  ) %>%
  tab_spanner(
    label = "Absolute Risk",
    columns = c(Control, Treatment)
  ) %>%
  cols_align("center", columns = everything()) %>%
  opt_row_striping() %>%
  tab_options(
    table.font.size = px(12),
    table.width = pct(100)
  ) %>%
  tab_source_note(
    source_note = md("*Incidence Rate Ratio instead of Odds Ratio")
  )

# Save as PNG 
gtsave(final, filename = "summary_table.png")


```

\newpage

## Subgroup Analyses - by Viruses

Models are re-run with COVID or Influenza alone.

```{r}
#| echo: false
#| warning: false

#run models for influenza alone
model_death_logistic_agg_f <- glmer(
  death ~ allocation_f + (1 | home),
  family = binomial(link = "logit"),
  data = subset(deaths_ob, virus=="FLU")
)

model_transfer_logistic_agg_f <- glmer(
  transfer ~ allocation_f + (1 | home),
  family = binomial(link = "logit"),
  data = subset(transfers_ob, virus=="FLU")
)

m1_f <- glmer(n ~ allocation_f +
                   (1|home), 
                 data=subset(out_num, virus=="FLU"), family = "poisson")

m2_f <- glmer (cbind(outbreak_size_secondary, noncases) ~ 
               allocation_f + (1|home), 
             family="binomial",
             data=subset(outbreak_size_model_y, virus=="FLU"))

m3_f <- glmer(
  cbind(outbreak_size_secondary, outbreak_size - outbreak_size_secondary) ~ allocation_f + (1|home),
  family = "binomial",
  data = subset(outbreak_size_model_y, virus=="FLU")
)

# Source the helpers
source("model_summary_helpers.R")

# Define models with their outcome labels
binary_models <- list(
  "Death" = model_death_logistic_agg_f,
  "Transfer" = model_transfer_logistic_agg_f,
  "Secondary Transmission" = m2_f,
  "Proportion of Secondary Cases per Outbreak" = m3_f
)

# Define count models with newdata for predictions
count_models_f <- list(
  list(model = m1_f, label = "Outbreaks", newdata = datagrid(model = m1, allocation_f = c("Control", "Intervention")))
)

# Create summary for all binary models
binary_summaries_f <- lapply(names(binary_models), function(label) {
  process_binary_model(binary_models[[label]], label)
})

# Create summary for all count models
count_summaries_f <- lapply(count_models_f, function(x) {
  process_count_model(x$model, x$label, newdata = x$newdata)
})

#create summary for outbreak size
outbreak_summaries_f <- process_outbreak_size_model(
  model = m2_f,
  data = outbreak_size_model_y,
  outcome_label = "Predicted Outbreak Size"
)

# Combine all results
final_table_all_f <- bind_rows(binary_summaries_f, count_summaries_f)
final_table_all_f <- bind_rows(final_table_all_f, outbreak_summaries_f)
```

```{r}
#| echo: false
#| warning: false

#run models for covid alone
model_death_logistic_agg_f <- glmer(
  death ~ allocation_f + (1 | home),
  family = binomial(link = "logit"),
  data = subset(deaths_ob, virus=="COVID")
)

model_transfer_logistic_agg_f <- glmer(
  transfer ~ allocation_f + (1 | home),
  family = binomial(link = "logit"),
  data = subset(transfers_ob, virus=="COVID")
)

m1_f <- glmer(n ~ allocation_f +
                   (1|home), 
                 data=subset(out_num, virus=="COVID"), family = "poisson")

m2_f <- glmer (cbind(outbreak_size_secondary, noncases) ~ 
               allocation_f + (1|home), 
             family="binomial",
             data=subset(outbreak_size_model, virus=="COVID"))

m3_f <- glmer(
  cbind(outbreak_size_secondary, outbreak_size - outbreak_size_secondary) ~ allocation_f + (1|home),
  family = "binomial",
  data = subset(outbreak_size_model_y, virus=="COVID")
)


# Source the helpers
source("model_summary_helpers.R")

# Define models with their outcome labels
binary_models <- list(
  "Death" = model_death_logistic_agg_f,
  "Transfer" = model_transfer_logistic_agg_f,
  "Secondary Transmission" = m2_f,
  "Proportion of Secondary Cases per Outbreak" = m3_f
)

# Define count models with newdata for predictions
count_models_f <- list(
  list(model = m1_f, label = "Outbreaks", newdata = datagrid(model = m1, allocation_f = c("Control", "Intervention")))
)

# Create summary for all binary models
binary_summaries_f <- lapply(names(binary_models), function(label) {
  process_binary_model(binary_models[[label]], label)
})

# Create summary for all count models
count_summaries_f <- lapply(count_models_f, function(x) {
  process_count_model(x$model, x$label, newdata = x$newdata)
})

#create summary for outbreak size
outbreak_summaries_f <- process_outbreak_size_model(
  model = m2_f,
  data = outbreak_size_model_y,
  outcome_label = "Predicted Outbreak Size"
)

# Combine all results
final_table_all_covid <- bind_rows(binary_summaries_f, count_summaries_f)
final_table_all_covid <- bind_rows(final_table_all_covid, outbreak_summaries_f)
```

\newpage

## Overall Transfer/Death Rates

```{r}
#| echo: false
#| warning: false

#create master database of all outcomes from all outcomes over study period
db_big <- left_join(db_check, db_final) %>%
  select(dummy_id, home, unit, outcome_date, combined_outcome, case_outcome, study_id,
         case_hosp_transfer_date, case_death_date, virus, outbreak_num) %>%
  mutate(suspect_case = case_when(is.na(study_id) ~ 1,
                                  TRUE ~ 0)) 
#add in beds 
random <- read_excel("prompt_ltc_home_unit_key_scrambled_07april2025.xlsx") %>%
  select(-allocation_f) %>%
  right_join(
    read_excel("randomization_key_08jan2025.xlsx"),
    by = "home"
  ) %>%
  right_join(
    read_excel("prompt_ltc_home_key_crowd_oh.xlsx"),
    by = "home"
  ) %>%
  mutate(
    allocation_f = case_when(
      allocation == "Platform" ~ "Intervention",
      allocation == "Control" ~ "Control"
    )
  ) %>%
  group_by(home, allocation, allocation_f) %>%
  dplyr::summarize(
    beds = sum(beds)
  ) %>%
  select(home, beds, allocation, allocation_f)
  
db_big <- left_join(db_big, random) 
  
#want rates per month per beds
# Step 1: Ensure date format and extract month
db_big <- db_big %>%
  mutate(
    outcome_date = as.Date(outcome_date),
    month = floor_date(outcome_date, "month")
  )

#Need beds per home
beds_per_home <- read_excel("prompt_ltc_home_unit_key_scrambled_07april2025.xlsx") %>%
  select(-allocation_f) %>%
  right_join(read_excel("randomization_key_08jan2025.xlsx") %>%
  right_join(read_excel("prompt_ltc_home_key_crowd_oh.xlsx")) %>%
  mutate(allocation_f = case_when(allocation=="Platform" ~ "Intervention",
                                  allocation=="Control" ~ "Control"))) %>%
  select(allocation_f, home, beds) %>%
  group_by(allocation_f) %>%
  dplyr::summarize(
    beds = sum(beds, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Aggregate outcome counts per group
outcome_counts <- db_big %>%
  filter(!is.na(combined_outcome)) %>%
  group_by(month, allocation_f, combined_outcome) %>%
  summarize(
    n = n()
  )

# Step 3: Create complete grid of all combinations to fill 0s
all_combinations <- crossing(
  month = seq(min(db_big$month, na.rm = TRUE), max(db_big$month, na.rm = TRUE), by = "month"),
  allocation_f = unique(db_big$allocation_f),
  combined_outcome = unique(db_big$combined_outcome)
)

# Step 4: Merge and compute rate per 100 beds
rates_df <- all_combinations %>%
  left_join(outcome_counts, by = c("month", "allocation_f", "combined_outcome")) %>%
  left_join(beds_per_home, by = "allocation_f") %>%
  mutate(
    n = replace_na(n, 0),
    rate_per_bed = (n / beds) * 100
  ) %>%
  select(month, allocation_f, combined_outcome, n, beds, rate_per_bed)

# Step 5: Filter out invalid or placeholder allocations if needed
rates_df <- rates_df %>%
  filter(!combined_outcome %in% c("", NA))
  
#plot monthly rates of transfers/deaths between allocation groups
library(wesanderson)
pp_1 <- rates_df %>%
  filter(month<"2025-05-01") %>%
  ggplot(aes(x = month, y = rate_per_bed, fill = combined_outcome)) +
  geom_bar(stat = "identity") +
  labs(
    #title = "Monthly Rate per Bed by Combined Outcome",
    x = "Month",
    y = "Rate of resident hospital transfer or death per 100 Beds",
    fill = "Outcome"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%b %Y"
  ) +
  #scale_fill_manual(values = wesanderson::wes_palette("Royal1", n = length(unique(rates_df$combined_outcome)), type = "discrete")) +
  scale_fill_grey(start = 0.3, end = 0.8) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  facet_grid(allocation_f~.)
```